WSClean: an implementation of a fast, generic wide-field imager

### 2014-02-28 articles/2014/Article-WSClean/wsclean.tex
Slight updates to formulae

For the WSClean time-cost function $t_\textrm{WSClean}$ we find 
\begin{align}\label{eq:wsclean-computing-cost}
& t_\textrm{WSClean} (w_{\max},N_\textrm{Mvis},N_\textrm{freq},N_\textrm{kpix}) = \\ \notag%
& N_\textrm{freq} \left( 0.452 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} (w_{\max} + 0.611) + 244 N_\textrm{Mvis} \right)
\end{align}
and for CASA we find
\begin{align}\label{eq:casa-computing-cost}
& t_\textrm{CASA}(w_{\max},N_\textrm{vis},N_\textrm{freq},N_\textrm{pix}) = \\ \notag%
& N_\textrm{freq} \left(1.03 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} + 5.93 N_\textrm{Mvis} (w_{\max} + 1.71)^2\right) + 172.
\label{eq:casa-computing-cost}%
\end{align}
Parameter $w_{\max}$ is the maximum $w$-value and is estimated with
\begin{equation}\resizebox{0.45\textwidth}{!}{$
 w_{\max} = \frac{1}{\lambda}\left[(D \sin \textrm{ZA} + \Delta z_{\max} \cos \textrm{ZA} + \xi) \left(1.0 - \cos(\frac{1}{2}\textrm{FOV})\right)\right]$},
\end{equation}%
where $D$ is the maximum baseline length, $\Delta z_{\max}$ is the maximum height difference between antennas and $\lambda$ is the wavelength, all in meters, and $\xi$ is an extra parameter for fitting the CASA measurements, $\xi_\textrm{CASA}=28.4$ and $\xi_\textrm{WSClean}=0$. These functions follow the trend of the measurements well, with an absolute error of 14.8\% and 20.7\% for respectively the WSClean and CASA functions.

@@

So far, we have corrected the (full-polarization) beam in image space, which is possible because of the homogeneity of the MWA tiles. This is cheap for our current snapshot time of 112s, but this might be infeasible when it is required to calculate the beam on very short time scales. Accurate MWA beam models are still being developed. When beam corrections are required on short time scales, calculating $a$-projection kernels or performing snapshot imaging also becomes more expensive, and it would be interesting to find out where the balance lies between these methods. When only a few thousand components need to be deconvolved, an approach with direct Fourier transforms is accurate and affordable. For wide-field arrays this can be combined with calibration of the direction dependence, e.g. with the SAGECAL \citep{sage-calibration-ii} or RTS peeling \citep{rts-mwa} calibration techniques. This is not possible for fields with diffuse emission or faint point sources when no model is known a-priory.

>>>>

For the WSClean time-cost function $t_\textrm{WSClean}$ we find 
\begin{align}\label{eq:wsclean-computing-cost}
& t_\textrm{WSClean} (w_{\max},N_\textrm{Mvis},N_\textrm{freq},N_\textrm{kpix}) = \\ \notag%
& N_\textrm{freq} \left( 0.526 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} (w_{\max} + 0.715) + 0.535 \right) + 0.248 N_\textrm{Mvis}
\end{align}
and for CASA we find
\begin{align}\label{eq:casa-computing-cost}
& t_\textrm{CASA}(w_{\max},N_\textrm{vis},N_\textrm{freq},N_\textrm{pix}) = \\ \notag%
& N_\textrm{freq} \left(0.965 N_\textrm{kpix}^2 \log_2 N_\textrm{kpix} + 0.0106 N_\textrm{Mvis} (w_{\max}^2 + 40.1)\right) + 40.8.
\label{eq:casa-computing-cost}%
\end{align}
Parameter $w_{\max}$ is the maximum $w$-value and is estimated with
\begin{equation}\resizebox{0.45\textwidth}{!}{$
 w_{\max} = \frac{1}{\lambda}\left[(D \sin \textrm{ZA} + \Delta z_{\max} \cos \textrm{ZA} + \xi) \left(1.0 - \cos(\frac{1}{2}\textrm{FOV})\right)\right]$},
\end{equation}%
where $D$ is the maximum baseline length, $\Delta z_{\max}$ is the maximum height difference between antennas and $\lambda$ is the wavelength, all in meters, and $\xi$ is an extra parameter for fitting the CASA measurements, $\xi_\textrm{CASA}=28.4$ and $\xi_\textrm{WSClean}=0$. These functions follow the trend of the measurements well, with an absolute error of 14.8\% and 20.7\% for respectively the WSClean and CASA functions.

@@

So far, we have corrected the (full-polarization) beam in image space, which is possible because of the homogeneity of the MWA tiles. This is cheap for our current snapshot time of 112s, but this might be infeasible when it is required to calculate the beam on very short time scales. Accurate MWA beam models are still being developed. When beam corrections are required on short time scales, calculating $a$-projection kernels or performing snapshot imaging also becomes more expensive, and it would be interesting to find out where the balance lies between these methods. When only a few thousand components need to be deconvolved, an approach with direct Fourier transforms is accurate and affordable. For wide-field arrays this can be combined with calibration of the direction dependence, e.g. with the SAGECAL \citep{sage-calibration-ii} or RTS peeling \citep{rts-mwa} calibration techniques. This is not possible for fields with diffuse emission or faint point sources when no model is known \textit{a priori}.
